---
title: "Statistical Modeling and Regression Project"
author: "Kushal"
date: "4/27/2023"
output: html_document
---

Reading data from january 1st 2015 until 2023-04-05. S&P 500 is considered for the whole process. I abstracted the data at a real time from yahoofinance using quantmod library..In the below code, initially we read stock data from yahoo finance calling the yahoofinance api through quantmod package and create a dataframe df.
```{r}

library(tidyverse)
library(quantmod)

begining <- as.Date("2015-01-01")
last_date <- as.Date("2023-04-05")
raw <- getSymbols("^GSPC", src = "yahoo", from = begining, to = last_date, auto.assign = FALSE)
stock_date = index(raw)
closingprice = as.numeric(raw[, "GSPC.Close"])
df <- data.frame(stock_date, closingprice)
```

#Using Linear Regression Model
```{r}
linear_model <- lm(closingprice ~ stock_date, data = df)
predict_stock <- predict(linear_model, newdata = df)
mse <- mean((predict_stock - df$closingprice)^2)
cat("MSE: ",mse,"\n")
rmse <- sqrt(mse)
cat("RMSE: ",rmse)
summary(linear_model)
```
```{r}
plot(linear_model) 
```


```{r}
predict_stock <- predict(linear_model, newdata = df)
ggplot() +
  geom_line(data = df, aes(x = stock_date, y = closingprice), color = "yellow") +
  geom_line(data = df, aes(x = stock_date, y = predict_stock), color = "red") +
  labs(title = "Prediction with Simple Linear Regression", x = "Stock Date", y = "Closing Price") 

```
#Using Random Forest Regression
```{r}
library(randomForest)
rand_model <- randomForest(closingprice ~ stock_date, data = df, ntree = 200, mtry = 1)
predict_stock <- predict(rand_model, newdata = df)
mse <- mean((predict_stock - df$closingprice)^2)
cat("MSE: ",mse,"\n")
rmse <- sqrt(mse)
cat("RMSE: ",rmse)
summary(rand_model)
```
```{r}
plot(rand_model) 
```

```{r}

ggplot() +
  geom_line(data = df, aes(x = stock_date, y = closingprice), color = "blue") +
  geom_line(data = df, aes(x = stock_date, y = predict_stock), color = "red") +
  labs(title = "Prediction with Random Forest Regression", x = "Stock Date", y = "Closing Price") 

```

#Using polynomial regression
```{r}
date_num <- as.numeric(df$stock_date)
poly_model <- lm(closingprice ~ poly(date_num, 2, raw = TRUE), data = df)
predict_stock <- predict(poly_model, newdata = df)
mse <- mean((predict_stock - df$closingprice)^2)
cat("MSE: ",mse, "\n")
rmse <- sqrt(mse)
cat("RMSE: ",rmse)
summary(poly_model)
```

```{r}
plot(poly_model) 

```

```{r}

ggplot() +
  geom_line(data = df, aes(x = stock_date, y = closingprice), color = "blue") +
  geom_line(data = df, aes(x = stock_date, y = predict_stock), color = "red") +
  labs(title = "Prediction with Random Forest Regression", x = "Stock Date", y = "Closing Price") 

```

